---
title: Barrier effects on the spatial distribution of *Xylella fastidiosa* in Alicante,
  Spain
author:
- affiliation: Centre de Protecció Vegetal i Biotecnologia, Institut Valencià d'Investigacions
    Agràries (IVIA), 46113 Moncada, Spain.
  email: cendoya_mar@externos.gva.es
  name: Martina Cendoya
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


```{r}
library(INLA)
library(rgdal)
library(ggplot2)
library(viridis)
library(gridExtra)
library(raster)
library(ggspatial)
library(rgeos)
```


## Data


```{r results='hide'}
# Xylella fastidiosa occurrence data in the demarcated area of Alicante (Spain)

data_Xf <- read.table("./Data/data_Xf.txt")

# Coordinate reference system

newproj <- "+proj=utm +zone=30 +ellps=GRS80 +units=m +no_defs"

```

```{r results='hide'}
## -- Boundaries -- ##

# Study area

bound <- readOGR('./Data/Boundaries', 'bound')

# Mountain barrier boundary

MB <- readOGR('./Data/Boundaries', 'bound_MB')

bound_MB <- gDifference(bound, MB, checkValidity = 2L)

# Continuous barrier boundary

CB <- readOGR('./Data/Boundaries', 'bound_CB')

bound_CB <- gDifference(bound, CB, checkValidity = 2L)

# Discontinuous barrier boundary

DB <- readOGR('./Data/Boundaries', 'bound_DB')

bound_DB <- gDifference(bound, DB, checkValidity = 2L)

```




```{r eval=FALSE}
## -- Remove data inside the barrier -- ##

# Mountain barrier

out.barrier = over(bound_MB,
                   SpatialPoints(cbind(data_Xf$X, data_Xf$Y), proj4string = CRS(newproj)),
                   returnList = T)[[1]]

data_M <- data_Xf[out.barrier,]

# write.table(data_M, './Data/data_Xf_M.txt')

# Perimeter barriers

out.barrier = over(bound_CB,
                   SpatialPoints(cbind(data_Xf$X, data_Xf$Y), proj4string = CRS(newproj)),
                   returnList = T)[[1]]

data_B <- data_Xf[out.barrier,]

# write.table(data_B, './Data/data_Xf_B.txt')

```


```{r results='hide'}
## -- Read data -- ##

# Mountain barrier data

data_M <- read.table("./Data/data_Xf_M.txt")

# Perimeter barriers data

data_B <- read.table("./Data/data_Xf_B.txt")

```


```{r echo=FALSE}
par(mai = c(0, 0, 0.5, 0), mfrow = c(2, 2), cex.main=0.8)

plot(bound, main = "Stationary")
points(data_Xf$X[data_Xf$ResLab == 0],
       data_Xf$Y[data_Xf$ResLab == 0],
       col = "green4",
       cex = 0.3,
       pch = 16)
points(data_Xf$X[data_Xf$ResLab == 1],
       data_Xf$Y[data_Xf$ResLab == 1],
       col = "red2",
       cex = 0.3,
       pch = 16)


plot(bound_MB, main="Mountain barrier")
plot(MB, add = T, col = "grey")
points(data_M$X[data_M$ResLab == 0],
       data_M$Y[data_M$ResLab == 0],
       col = "green4",
       cex = 0.3,
       pch = 16)
points(data_M$X[data_M$ResLab == 1],
       data_M$Y[data_M$ResLab == 1],
       col = "red2",
       cex = 0.3,
       pch = 16)


plot(bound_CB, main="Continuous barrier")
plot(CB, add = T, col = "grey")
points(data_B$X[data_B$ResLab == 0],
       data_B$Y[data_B$ResLab == 0],
       col = "green4",
       cex = 0.3,
       pch = 16)
points(data_B$X[data_B$ResLab == 1],
       data_B$Y[data_B$ResLab == 1],
       col = "red2",
       cex = 0.3,
       pch = 16)


plot(bound_DB, main="Discontinuous barrier")
plot(DB, add = T, col = "grey")
points(data_B$X[data_B$ResLab == 0],
       data_B$Y[data_B$ResLab == 0],
       col = "green4",
       cex = 0.3,
       pch = 16)
points(data_B$X[data_B$ResLab == 1],
       data_B$Y[data_B$ResLab == 1],
       col = "red2",
       cex = 0.3,
       pch = 16)

```

## Models

```{r}
## -- Mesh -- ##

# Stationary

mesh_ST <- inla.mesh.2d(
  boundary = bound,
  max.edge = c(750, 3500),
  offset = c(0.1,-0.1),
  cutoff   = 750
)

# Mountain barrier

mesh_MB <- inla.mesh.2d(
  boundary = bound_MB,
  max.edge = c(750, 3500),
  offset = c(0.1, -0.1),
  cutoff   = 750
)


# Continuous barrier

mesh_CB <- inla.mesh.2d(
  boundary = bound_CB,
  max.edge = c(750, 3500),
  offset = c(0.1, -0.1),
  cutoff   = 750
)

# Discontinuous barrier

mesh_DB <- inla.mesh.2d(
  boundary = bound_DB,
  max.edge = c(750, 3500),
  offset = c(0.1, -0.1),
  cutoff   = 750
)
```


```{r echo=FALSE}
# Plot

par(mai = c(0, 0, 0, 0), mfrow = c(2, 2))

plot(mesh_ST, main = F, lwd = 0.5)

plot(mesh_MB, main = F, lwd = 0.5)
plot(MB, add = T, col = "grey")

plot(mesh_CB, main = F, lwd = 0.5)
plot(CB, add = T, col = "grey")

plot(mesh_DB, main = F, lwd = 0.5)
plot(DB, add = T, col = "grey")

```


```{r include=FALSE}
# Function to convert matrix to raster

rotate <- function(x)
   {apply(x, 2, rev)}

matrix_to_raster <- function(m, proj.grid.mat = proj.grid.mat)
{
  raster(
    rotate(t(m)),
    xmn = min(proj.grid.mat$lattice$loc[, 1]),
    xmx = max(proj.grid.mat$lattice$loc[, 1]),
    ymn = min(proj.grid.mat$lattice$loc[, 2]),
    ymx = max(proj.grid.mat$lattice$loc[, 2]),
    crs = proj4string(bound)
  )
}

```

### 1. Stationary model

```{r }
data <- data_Xf
mesh <- mesh_ST
```

```{r}
# SPDE model definition

size <- min(c(diff(range(data$X)), diff(range(data$Y))))
range0 <- size / 2

spde <- inla.spde2.pcmatern(
  # Mesh
  mesh = mesh,
  # P(practic.range < range0) = 0.5
  prior.range = c(range0, 0.5),
  # P(sigma > 10) = 0.01
  prior.sigma = c(10, 0.01)
)

# Projector matrix

A.est <- inla.spde.make.A(mesh, loc = cbind(data$X, data$Y))

# Data stack

stk.est <- inla.stack(
  data    = list(y = data$ResLab),
  A       = list(A.est, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, nrow(data))),
  tag     = 'est'
)
```

```{r eval=FALSE}
# Model fitting

formula.1 <- y ~ -1 + beta0 + f(spatial, model = spde)

model.est_ST <- inla(
  formula.1,
  data              = inla.stack.data(stk.est),
  family            = "binomial" ,
  control.compute   = list(
    dic              = TRUE,
    cpo              = TRUE,
    waic             = TRUE,
    return.marginals = TRUE
  ),
  control.predictor = list(A       = inla.stack.A(stk.est),
                           compute = TRUE),
  num.threads       = 2,
  verbose           = FALSE
)

saveRDS(model.est_ST, "./Models/ST_model_est.rds")

```

```{r}
# Read saved model

model.est_ST <- readRDS("./Models/ST_model_est.rds")
```


```{r}
# Projection on a grid

dxy <- apply(bbox(bound), 1, diff)
r <- dxy[1] / dxy[2]
m <- 120
proj.grid.mat <-
  inla.mesh.projector(
    mesh,
    xlim = bbox(bound)[1, ],
    ylim = bbox(bound)[2, ],
    dims = c(r, 1) * m
  )

# NA to the values outside boundary

ov <-
  over(
    SpatialPoints(proj.grid.mat$lattice$loc, bound@proj4string),
    SpatialPolygons(bound@polygons, proj4string = bound@proj4string)
  )

i.temp <- is.na(ov)

# Projector matrix to predict

A.pred <-
  inla.spde.make.A(mesh, loc = proj.grid.mat$lattice$loc[!i.temp, ])

# Stack to predict

stk.pred <- inla.stack(
  data      = list(y = NA),
  A       = list(A.pred, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, dim(A.pred)[1])),
  tag     = 'pred'
)

stk <- inla.stack(stk.est, stk.pred)

```

```{r eval=FALSE}
# Prediction

model.pred_ST <- inla(
  formula.1,
  data = inla.stack.data(stk),
  family = "binomial",
  control.predictor = list(
    A       = inla.stack.A(stk),
    compute = TRUE,
    link    = 1
  ),
  control.mode      = list(theta = model.est_ST$mode$theta,
                           restart = TRUE),
  num.threads       = 3,
  verbose  = TRUE
)

saveRDS(model.pred_ST, "./Models/ST_model_pred.rds")

```

```{r}
# Read saved model

model.pred_ST <- readRDS("./Models/ST_model_pred.rds")
```


```{r}
# Index of the predicted values

idx <- inla.stack.index(stk, 'pred')$data

# Matrix to visualize

prob.mean_ST <-
  prob.sd_ST <- matrix(NA, proj.grid.mat$lattice$dims[1],
                       proj.grid.mat$lattice$dims[2])
prob.mean_ST[!i.temp] <-
  c(model.pred_ST$summary.fitted.values$mean[idx])
prob.sd_ST[!i.temp] <-
  c(model.pred_ST$summary.fitted.values$sd[idx])
```

#### Results

- Posterior marginal distribution of parameters

```{r}
summary(model.est_ST)$fixed
```

&nbsp;

- Posterior marginal distribution of hyperparameters

```{r}
summary(model.est_ST)$hyperpar

```

&nbsp;

- Mean and SD of the posterior predictive distribution

```{r echo=FALSE, fig.width=15, fig.height=5}
# Plot mean and SD of the posterior predictive

df_ST <- expand.grid(x = proj.grid.mat$x, y = proj.grid.mat$y)
df_ST$mean_p <- as.vector(prob.mean_ST)
df_ST$sd_p <- as.vector(prob.sd_ST)

meanpred <- ggplot(df_ST, aes(x = x, y = y, fill = mean_p)) +
  geom_tile() +
  geom_polygon(
    data = bound,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_gradient2(
    low = "white",
    high = "red",
    na.value = "transparent",
    limits = c(0, 0.87)
  )  +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) + 
  ggtitle("Mean predictive")


sdpred <- ggplot(df_ST, aes(x = x, y = y, fill = sd_p)) +
  geom_tile() +
  geom_polygon(
    data = bound,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_viridis(
    option = "D",
    begin = 0.2,
    direction = -1,
    na.value = "transparent",
    limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.05)
  ) +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("SD predictive")


grid.arrange(meanpred, sdpred, ncol = 2)

```

&nbsp;

```{r}
# Raster of the posterior predictive mean for the difference between the models

mean_pred_ST <- matrix_to_raster(prob.mean_ST, proj.grid.mat = proj.grid.mat)
mean_pred_ST<-mask(mean_pred_ST, bound)
```


### 2. Mountain barrier model

```{r}
data <- data_M
bound_barriers <- bound_MB
mesh <- mesh_MB

# Polygon with the barrier

in.tri <- inla.over_sp_mesh(bound_barriers,
                            y = mesh,
                            type = "centroid",
                            ignore.CRS = TRUE)
num.tri <- length(mesh$graph$tv[, 1])
barrier.tri <- setdiff(1:num.tri, in.tri)
poly.barrier <- inla.barrier.polygon(mesh,
                                     barrier.triangles = barrier.tri)

# SPDE model definition

size <- min(c(diff(range(data$X)), diff(range(data$Y))))
range0 <- size / 2

barrier.model <- inla.barrier.pcmatern(
  mesh,
  barrier.triangles =
    barrier.tri,
  prior.range = c(range0, 0.5),
  prior.sigma = c(10, 0.01)
)


# Projector matrix

A.est <- inla.spde.make.A(mesh, loc = cbind(data$X, data$Y))

# Data stack

stk.est <- inla.stack(
  data    = list(y = data$ResLab),
  A       = list(A.est, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, nrow(data))),
  tag     = 'est'
)
```

```{r eval=FALSE}
# Model fitting

formula.1 <- y ~ -1 + beta0 + f(spatial, model = barrier.model)

model.est_MB <- inla(
  formula.1,
  data              = inla.stack.data(stk.est),
  family            = "binomial" ,
  control.compute   = list(
    dic              = TRUE,
    cpo              = TRUE,
    waic             = TRUE,
    return.marginals = TRUE
  ),
  control.predictor = list(A       = inla.stack.A(stk.est),
                           compute = TRUE),
  num.threads       = 2,
  verbose           = FALSE
)


saveRDS(model.est_MB, "./Models/MB_model_est.rds")
```

```{r}
# Read saved model

model.est_MB <- readRDS("./Models/MB_model_est.rds")
```


```{r}
# Projection on a grid

dxy <- apply(bbox(bound_barriers), 1, diff)
r <- dxy[1] / dxy[2]
m <- 120
proj.grid.mat <-
  inla.mesh.projector(
    mesh,
    xlim = bbox(bound_barriers)[1, ],
    ylim = bbox(bound_barriers)[2, ],
    dims = c(r, 1) * m
  )

# NA to the values outside boundary

ov <-
  over(
    SpatialPoints(proj.grid.mat$lattice$loc, bound@proj4string),
    SpatialPolygons(bound@polygons, proj4string = bound@proj4string)
  )

i.temp <- is.na(ov)

# Projector matrix to predict

A.pred <-
  inla.spde.make.A(mesh, loc = proj.grid.mat$lattice$loc[!i.temp, ])

# Stack to predict

stk.pred <- inla.stack(
  data      = list(y = NA),
  A       = list(A.pred, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, dim(A.pred)[1])),
  tag     = 'pred'
)

stk <- inla.stack(stk.est, stk.pred)

```

```{r eval=FALSE}
# Prediction

model.pred_MB <- inla(
  formula.1,
  data = inla.stack.data(stk),
  family = "binomial",
  control.predictor = list(
    A       = inla.stack.A(stk),
    compute = TRUE,
    link    = 1
  ),
  control.mode      = list(theta = model.est_MB$mode$theta,
                           restart = TRUE),
  num.threads       = 3,
  verbose  = TRUE
)

saveRDS(model.pred_MB, "./Models/MB_model_pred.rds")

```

```{r}
# Read saved model

model.pred_MB <- readRDS("./Models/MB_model_pred.rds")
```

```{r}
# Index of the predicted values

idx <- inla.stack.index(stk, 'pred')$data

# Matrix to visualize

prob.mean_MB <-
  prob.sd_MB <- matrix(NA, proj.grid.mat$lattice$dims[1],
                       proj.grid.mat$lattice$dims[2])
prob.mean_MB[!i.temp] <-
  c(model.pred_MB$summary.fitted.values$mean[idx])
prob.sd_MB[!i.temp] <-
  c(model.pred_MB$summary.fitted.values$sd[idx])

```

#### Results

- Posterior marginal distribution of parameters

```{r}
summary(model.est_MB)$fixed
```

&nbsp;

- Posterior marginal distribution of hyperparameters

```{r}
summary(model.est_MB)$hyperpar
```

&nbsp;

Standard deviation of the spatial effect $= e^{\theta_1}$

Range $= e^{\theta_2}$

```{r}
pander::pander(exp(model.est_MB$summary.hyperpar[,c(1,3,4,5)]), row.names=c("SD", "Range"))
```


&nbsp;


- Mean and SD of the posterior predictive distribution

```{r echo=FALSE, fig.width=15, fig.height=5}
# Plot mean and SD of the posterior predictive

df_MB <- expand.grid(x = proj.grid.mat$x, y = proj.grid.mat$y)
df_MB$mean_p <- as.vector(prob.mean_MB)
df_MB$sd_p <- as.vector(prob.sd_MB)

barrier_MB <- crop(MB, bound)

meanpred <- ggplot(df_MB, aes(x = x, y = y, fill = mean_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_MB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_MB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_gradient2(
    low = "white",
    high = "red",
    na.value = "transparent",
    limits = c(0, 0.87)
  )  +   theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("Mean predictive")

sdpred <- ggplot(df_MB, aes(x = x, y = y, fill = sd_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_MB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_MB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_viridis(
    option = "D",
    begin = 0.2,
    direction = -1,
    na.value = "transparent",
    limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.05)
  ) +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("SD predictive")


grid.arrange(meanpred, sdpred, ncol = 2)

```

&nbsp;


```{r}
# Raster of the posterior predictive mean for the difference between the models.

mean_pred_MB <-
  matrix_to_raster(prob.mean_MB, proj.grid.mat = proj.grid.mat)
mean_pred_MB <- mask(mean_pred_MB, bound)

```


### 3. Continuous barrier model

```{r}
data <- data_B
bound_barriers <- bound_CB
mesh <- mesh_CB

# Polygon with the barrier

in.tri <- inla.over_sp_mesh(bound_barriers,
                            y = mesh,
                            type = "centroid",
                            ignore.CRS = TRUE)
num.tri <- length(mesh$graph$tv[, 1])
barrier.tri <- setdiff(1:num.tri, in.tri)
poly.barrier <- inla.barrier.polygon(mesh,
                                     barrier.triangles = barrier.tri)

# SPDE model definition

size <- min(c(diff(range(data$X)), diff(range(data$Y))))
range0 <- size / 2

barrier.model <- inla.barrier.pcmatern(
  mesh,
  barrier.triangles =
    barrier.tri,
  prior.range = c(range0, 0.5),
  prior.sigma = c(10, 0.01)
)


# Projector matrix

A.est <- inla.spde.make.A(mesh, loc = cbind(data$X, data$Y))

# Data stack

stk.est <- inla.stack(
  data    = list(y = data$ResLab),
  A       = list(A.est, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, nrow(data))),
  tag     = 'est'
)
```

```{r eval=FALSE}
# Model fitting

formula.1 <- y ~ -1 + beta0 + f(spatial, model = barrier.model)

model.est_CB <- inla(
  formula.1,
  data              = inla.stack.data(stk.est),
  family            = "binomial" ,
  control.compute   = list(
    dic              = TRUE,
    cpo              = TRUE,
    waic             = TRUE,
    return.marginals = TRUE
  ),
  control.predictor = list(A       = inla.stack.A(stk.est),
                           compute = TRUE),
  num.threads       = 2,
  verbose           = FALSE
)


saveRDS(model.est_CB, "./Models/CB_model_est.rds")
```

```{r}
# Read saved model

model.est_CB <- readRDS("./Models/CB_model_est.rds")
```


```{r}
# Projection on a grid

dxy <- apply(bbox(bound_barriers), 1, diff)
r <- dxy[1] / dxy[2]
m <- 120
proj.grid.mat <-
  inla.mesh.projector(
    mesh,
    xlim = bbox(bound_barriers)[1, ],
    ylim = bbox(bound_barriers)[2, ],
    dims = c(r, 1) * m
  )

# NA to the values outside boundary

ov <-
  over(
    SpatialPoints(proj.grid.mat$lattice$loc, bound@proj4string),
    SpatialPolygons(bound@polygons, proj4string = bound@proj4string)
  )

i.temp <- is.na(ov)

# Projector matrix to predict

A.pred <-
  inla.spde.make.A(mesh, loc = proj.grid.mat$lattice$loc[!i.temp, ])

# Stack to predict

stk.pred <- inla.stack(
  data      = list(y = NA),
  A       = list(A.pred, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, dim(A.pred)[1])),
  tag     = 'pred'
)

stk <- inla.stack(stk.est, stk.pred)

```

```{r eval=FALSE}
# Prediction

model.pred_CB <- inla(
  formula.1,
  data = inla.stack.data(stk),
  family = "binomial",
  control.predictor = list(
    A       = inla.stack.A(stk),
    compute = TRUE,
    link    = 1
  ),
  control.mode      = list(theta = model.est_CB$mode$theta,
                           restart = TRUE),
  num.threads       = 3,
  verbose  = TRUE
)

saveRDS(model.pred_CB, "./Models/CB_model_pred.rds")

```

```{r}
# Read saved model

model.pred_CB <- readRDS("./Models/CB_model_pred.rds")
```

```{r}
# Index of the predicted values

idx <- inla.stack.index(stk, 'pred')$data

# Matrix to visualize

prob.mean_CB <-
  prob.sd_CB <- matrix(NA, proj.grid.mat$lattice$dims[1],
                       proj.grid.mat$lattice$dims[2])
prob.mean_CB[!i.temp] <-
  c(model.pred_CB$summary.fitted.values$mean[idx])
prob.sd_CB[!i.temp] <-
  c(model.pred_CB$summary.fitted.values$sd[idx])

```

#### Results

- Posterior marginal distribution of parameters

```{r}
summary(model.est_CB)$fixed
```

&nbsp;

- Posterior marginal distribution of hyperparameters

```{r}
summary(model.est_CB)$hyperpar
```

&nbsp;

Standard deviation of the spatial effect $= e^{\theta_1}$

Range $= e^{\theta_2}$

```{r}
pander::pander(exp(model.est_CB$summary.hyperpar[,c(1,3,4,5)]), row.names=c("SD", "Range"))
```


&nbsp;


- Mean and SD of the posterior predictive distribution

```{r echo=FALSE, fig.width=15, fig.height=5}
# Plot mean and SD of the posterior predictive

df_CB <- expand.grid(x = proj.grid.mat$x, y = proj.grid.mat$y)
df_CB$mean_p <- as.vector(prob.mean_CB)
df_CB$sd_p <- as.vector(prob.sd_CB)

barrier_CB <- crop(CB, bound)

meanpred <- ggplot(df_CB, aes(x = x, y = y, fill = mean_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_CB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_CB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_gradient2(
    low = "white",
    high = "red",
    na.value = "transparent",
    limits = c(0, 0.87)
  )  +   theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("Mean predictive")

sdpred <- ggplot(df_CB, aes(x = x, y = y, fill = sd_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_CB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_CB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_viridis(
    option = "D",
    begin = 0.2,
    direction = -1,
    na.value = "transparent",
    limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.05)
  ) +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("SD predictive")


grid.arrange(meanpred, sdpred, ncol = 2)

```

```{r}
# Raster of the posterior predictive mean for the difference between the models.

mean_pred_CB <-
  matrix_to_raster(prob.mean_CB, proj.grid.mat = proj.grid.mat)
mean_pred_CB <- mask(mean_pred_CB, bound)

```

### 4. Discontinuous barrier model

```{r}
data <- data_B
bound_barriers <- bound_DB
mesh <- mesh_DB

# Polygon with the barrier

in.tri <- inla.over_sp_mesh(bound_barriers,
                            y = mesh,
                            type = "centroid",
                            ignore.CRS = TRUE)
num.tri <- length(mesh$graph$tv[, 1])
barrier.tri <- setdiff(1:num.tri, in.tri)
poly.barrier <- inla.barrier.polygon(mesh,
                                     barrier.triangles = barrier.tri)

# SPDE model definition

size <- min(c(diff(range(data$X)), diff(range(data$Y))))
range0 <- size / 2

barrier.model <- inla.barrier.pcmatern(
  mesh,
  barrier.triangles =
    barrier.tri,
  prior.range = c(range0, 0.5),
  prior.sigma = c(10, 0.01)
)


# Projector matrix

A.est <- inla.spde.make.A(mesh, loc = cbind(data$X, data$Y))

# Data stack

stk.est <- inla.stack(
  data    = list(y = data$ResLab),
  A       = list(A.est, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, nrow(data))),
  tag     = 'est'
)
```

```{r eval=FALSE}
# Model fitting

formula.1 <- y ~ -1 + beta0 + f(spatial, model = barrier.model)

model.est_DB <- inla(
  formula.1,
  data              = inla.stack.data(stk.est),
  family            = "binomial" ,
  control.compute   = list(
    dic              = TRUE,
    cpo              = TRUE,
    waic             = TRUE,
    return.marginals = TRUE
  ),
  control.predictor = list(A       = inla.stack.A(stk.est),
                           compute = TRUE),
  num.threads       = 2,
  verbose           = FALSE
)


saveRDS(model.est_DB, "./Models/DB_model_est.rds")
```

```{r}
# Read saved model

model.est_DB <- readRDS("./Models/DB_model_est.rds")
```


```{r}
# Projection on a grid

dxy <- apply(bbox(bound_barriers), 1, diff)
r <- dxy[1] / dxy[2]
m <- 120
proj.grid.mat <-
  inla.mesh.projector(
    mesh,
    xlim = bbox(bound_barriers)[1, ],
    ylim = bbox(bound_barriers)[2, ],
    dims = c(r, 1) * m
  )

# NA to the values outside boundary

ov <-
  over(
    SpatialPoints(proj.grid.mat$lattice$loc, bound@proj4string),
    SpatialPolygons(bound@polygons, proj4string = bound@proj4string)
  )

i.temp <- is.na(ov)

# Projector matrix to predict

A.pred <-
  inla.spde.make.A(mesh, loc = proj.grid.mat$lattice$loc[!i.temp, ])

# Stack to predict

stk.pred <- inla.stack(
  data      = list(y = NA),
  A       = list(A.pred, 1),
  effects = list(spatial = 1:mesh$n,
                 beta0 = rep(1, dim(A.pred)[1])),
  tag     = 'pred'
)

stk <- inla.stack(stk.est, stk.pred)

```

```{r eval=FALSE}
# Prediction

model.pred_DB <- inla(
  formula.1,
  data = inla.stack.data(stk),
  family = "binomial",
  control.predictor = list(
    A       = inla.stack.A(stk),
    compute = TRUE,
    link    = 1
  ),
  control.mode      = list(theta = model.est_DB$mode$theta,
                           restart = TRUE),
  num.threads       = 3,
  verbose  = TRUE
)

saveRDS(model.pred_DB, "./Models/DB_model_pred.rds")

```

```{r}
# Read saved model

model.pred_DB <- readRDS("./Models/DB_model_pred.rds")
```

```{r}
# Index of the predicted values

idx <- inla.stack.index(stk, 'pred')$data

# Matrix to visualize

prob.mean_DB <-
  prob.sd_DB <- matrix(NA, proj.grid.mat$lattice$dims[1],
                       proj.grid.mat$lattice$dims[2])
prob.mean_DB[!i.temp] <-
  c(model.pred_DB$summary.fitted.values$mean[idx])
prob.sd_DB[!i.temp] <-
  c(model.pred_DB$summary.fitted.values$sd[idx])

```

#### Results

- Posterior marginal distribution of parameters

```{r}
summary(model.est_CB)$fixed
```

&nbsp;

- Posterior marginal distribution of hyperparameters

```{r}
summary(model.est_CB)$hyperpar
```

&nbsp;

Standard deviation of the spatial effect $= e^{\theta_1}$

Range $= e^{\theta_2}$

```{r}
pander::pander(exp(model.est_CB$summary.hyperpar[,c(1,3,4,5)]), row.names=c("SD", "Range"))
```


&nbsp;


- Mean and SD of the posterior predictive distribution


```{r echo=FALSE, fig.width=15, fig.height=5}
# Plot mean and SD of the posterior predictive

df_DB <- expand.grid(x = proj.grid.mat$x, y = proj.grid.mat$y)
df_DB$mean_p <- as.vector(prob.mean_DB)
df_DB$sd_p <- as.vector(prob.sd_DB)

barrier_DB <- crop(DB, bound)

meanpred <- ggplot(df_DB, aes(x = x, y = y, fill = mean_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_DB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_DB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_gradient2(
    low = "white",
    high = "red",
    na.value = "transparent",
    limits = c(0, 0.87)
  )  +   theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("Mean predictive")

sdpred <- ggplot(df_DB, aes(x = x, y = y, fill = sd_p)) +
  geom_tile() +
  geom_polygon(
    data = barrier_DB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 1
  ) +
  geom_polygon(
    data = bound_DB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 1
  ) +
  scale_fill_viridis(
    option = "D",
    begin = 0.2,
    direction = -1,
    na.value = "transparent",
    limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.05)
  ) +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1.2, "cm"),
    legend.text = element_text(size = 15)
  ) +
  ggtitle("SD predictive")


grid.arrange(meanpred, sdpred, ncol = 2)

```

```{r}
# Raster of the posterior predictive mean for the difference between the models.

mean_pred_DB <-
  matrix_to_raster(prob.mean_DB, proj.grid.mat = proj.grid.mat)
mean_pred_DB <- mask(mean_pred_DB, bound)

```

## Differences of the posterior predictive mean

```{r include=FALSE}
# Stationary - Continuous barrier

preddiff_STCB <- overlay(
  mean_pred_ST,
  mean_pred_CB,
  fun = function(r1, r2) {
    return(r1 - r2)
  }
)

# Stationary - Discontinuous barrier

preddiff_STDB <- overlay(
  mean_pred_ST,
  mean_pred_DB,
  fun = function(r1, r2) {
    return(r1 - r2)
  }
)

# Stationary - Mountain barrier

preddiff_DBCB <- overlay(
  mean_pred_DB,
  mean_pred_CB,
  fun = function(r1, r2) {
    return(r1 - r2)
  }
)

# Discontinuous barrier - Continuous barrier

preddiff_STMB <- overlay(
  mean_pred_ST,
  mean_pred_MB,
  fun = function(r1, r2) {
    return(r1 - r2)
  }
)

# Raster to points

preddiff_STCB_pts <- rasterToPoints(preddiff_STCB, spatial = TRUE)
preddiff_STDB_pts <- rasterToPoints(preddiff_STDB, spatial = TRUE)
preddiff_DBCB_pts <- rasterToPoints(preddiff_DBCB, spatial = TRUE)
preddiff_STMB_pts <- rasterToPoints(preddiff_STMB, spatial = TRUE)

# Data frame

df_diffpred <-
  data.frame(x = preddiff_STCB_pts@coords[, 1], y = preddiff_STCB_pts@coords[, 2])
df_diffpred$STCB <- preddiff_STCB_pts@data$layer
df_diffpred$STDB <- preddiff_STDB_pts@data$layer
df_diffpred$DBCB <- preddiff_DBCB_pts@data$layer
df_diffpred$STMB <- preddiff_STMB_pts@data$layer

```


```{r echo=FALSE}
# Plot differences

pred_STCB <- ggplot(df_diffpred, aes(x = x, y = y, fill = STCB)) +
  geom_tile() +
  geom_polygon(
    data = barrier_CB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 0.8
  ) +
  geom_polygon(
    data = bound_CB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 0.8
  ) +
  scale_fill_gradient2(
    midpoint = 0,
    low = "blue",
    mid = "white",
    high = "red",
    na.value = "transparent",
    limits = c(-0.48, 0.34)
  )  +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1, "cm"),
    legend.text = element_text(size = 8),
    plot.title = element_text(size=8)
  ) +
  ggtitle("Difference mean predictive \nStationary - Continuous barrier")

pred_STDB <- ggplot(df_diffpred, aes(x = x, y = y, fill = STDB)) +
  geom_tile() +
  geom_polygon(
    data = barrier_DB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 0.8
  ) +
  geom_polygon(
    data = bound_CB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 0.8
  ) +
  scale_fill_gradient2(
    midpoint = 0,
    low = "blue",
    mid = "white",
    high = "red",
    na.value = "transparent",
    limits=c(-0.48, 0.34)
  )  +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1, "cm"),
    legend.text = element_text(size = 8), 
    plot.title = element_text(size=8)
  ) +
  ggtitle("Difference mean predictive \nStationary - Discontinuous barrier")

pred_DBCB <- ggplot(df_diffpred, aes(x = x, y = y, fill = DBCB)) +
  geom_tile() +
  geom_polygon(
    data = barrier_DB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 0.8
  ) +
  geom_polygon(
    data = bound_CB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 0.8
  ) +
  scale_fill_gradient2(
    midpoint = 0,
    low = "blue",
    mid = "white",
    high = "red",
    na.value = "transparent",
    limits=c(-0.48, 0.34)
  )  +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1, "cm"),
    legend.text = element_text(size = 8), 
    plot.title = element_text(size=8)
  ) +
  ggtitle("Difference mean predictive \nDiscontinuous barrier - Continuous barrier")

pred_STMB <- ggplot(df_diffpred, aes(x = x, y = y, fill = STMB)) +
  geom_tile() +
  geom_polygon(
    data = barrier_MB,
    aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "grey50",
    size = 0.8
  ) +
  geom_polygon(
    data = bound_MB,
    aes(x = long, y = lat, group = group),
    fill = NA,
    color = "grey50",
    size = 0.8
  ) +
  scale_fill_gradient2(
    midpoint = 0,
    low = "blue",
    mid = "white",
    high = "red",
    na.value = "transparent",
    limits=c(-0.48, 0.34)
  )  +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(1, "cm"),
    legend.text = element_text(size = 8), 
    plot.title = element_text(size=8)
  ) +
  ggtitle("Difference mean predictive \nStationary - Mountain barrier")


grid.arrange(pred_STCB, pred_STDB, pred_DBCB, pred_STMB, ncol = 2)
```
